---
title: "lab_1_in_class"
author: "Priscilla Hare"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objectives:

- basic review of data wrangling
- intro to spatial data in R with sf
- into to Shiny apps

1. Attatch packages
```{r}

library(tidyverse)
library(sf)
library(leaflet)
library(tmap)

```

2. Review on reading, exploring, wrangling, and visualizing data

a. Read in data
```{r}

wine <- read_csv("wine_data.csv")

```

b. Wrangling
```{r}

#only keep columns country, province, winery, region_1, points, and price
#rename province to state and then only keep Oregon, Washington, and California
#create a new column for points per dollar value (ppd) and arrange high to low
wine_new <- wine %>%
  select(country, province, winery, region_1, points, price) %>% 
  rename(state = province) %>% 
  filter(state == "California"|
         state == "Oregon"|
         state == "Washington") %>% 
  mutate(ppd = points/price) %>% 
  arrange(-ppd)

```

c. Use group_by() and summarize() to find summary statistics by group
```{r}

wine_summary <- wine_new %>% 
  group_by(state) %>% 
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE), #na.rm=TRUE removes 'NA' values before calculation
    max_ppd = max(ppd, na.rm = TRUE)
  )

```

d. A few gg-graphics

Histrograms
```{r}

#side by side histograms
wine_hists <- ggplot(wine_new, aes(x = ppd)) +
  geom_histogram(aes(fill = state)) +
  facet_wrap(~state, scales = "free")

wine_hists


```

Scatterplot graph of price vs. points, with color determined by state, adding a linear model trendline on top of it
```{r}

wine_points <- ggplot(wine_new, aes(x = price, y = points)) +
  geom_point(aes(color = state), alpha = 0.5) + #aalpha = 0.5 changes transparency
  geom_smooth(color = "black", method = "lm", se = FALSE) +
  scale_color_manual(values = c("royalblue", "mediumorchid1", "orange")) +
  facet_wrap(~state) +
  theme_light() 

wine_points
```

2. Intro to spatial data in R: California Counties spatial data using sf

a. Read in the data
```{r}

ca_counties <- st_read(dsn = ".", layer = "CA_Counties_TIGER2016") #use st_read to read in spatial data

#plot(ca_counties)

#simplify to only include county name and area of land

ca_land <- ca_counties %>% 
  select(NAME, ALAND)

ca_pop_inc <- read_csv("ca_pop_inc.csv") %>% 
  rename(NAME = COUNTY)

#merge spatial data with population csv data
ca_df <- full_join(ca_land, ca_pop_inc) %>% 
  select(NAME, MedFamilyIncome)

```
d. Use geom_sf to make a map in ggplot
```{r}

ca_income <- ggplot(ca_df) +
  geom_sf(aes(fill = MedFamilyIncome))

ca_income


ca_df_transform <- st_transform(ca_df, crs = 4326)

leaflet(ca_df_transform) %>% 
  addTiles() %>% 
  addPolygons()

```

